---
title: "Advanced Markdown: Dynamic Equations"
author: "Tilburg Science Hub"


# The next lines allow you to change the spacing in your document You can 
# switch out \onehalfspacing with \singlespacing or \doublespacing, if desired.
header-includes:
    - \linespread{1.25}
    - \usepackage[tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in]{geometry}
    - \usepackage{titlesec}
    - \titleformat{\chapter}[display]{\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{16pt}{\Large}
    - \titlespacing*{\chapter}{0pt}{-20pt}{40pt}
    
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output: 
  # bookdown::pdf_document2
  bookdown::html_document2
  # bookdown::word_document2
  # bookdown::gitbook

# Specify the location of the bibliography below
bibliography: references.bib
link-citations: yes

# Incorporate a list of figures or/and tables by setting the value to TRUE.
lof: false
---

\begin{equation} 
\operatorname{Sepal.Length} = \alpha + \beta_{1}(\operatorname{Species}_{\operatorname{versicolor}}) + \beta_{2}(\operatorname{Species}_{\operatorname{virginica}}) + \epsilon
  (\#eq:regression-1)
\end{equation} 



# Overview {-} 
<!-- The {.unnumbered} option here means that the overview will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

[`Rmarkdown`](https://rmarkdown.rstudio.com/) is a format designed for crafting reproducible and dynamic reports using R. It's a useful tool for creating a wide range of academic documents, including journal articles, books, and technical reports. 

Our website offers the `tisemdown` thesis [template](https://tilburgsciencehub.com/topics/research-skills/templates-dynamic-content/templates/bookdown-theses/), which provides a read-to-go thesis template using `Rmarkdown`. This template pre-configures all the formatting requirements of your study programs, kickstarting the initial setup process. 

This article builds upon the thesis template article by exploring more sophisticated `Rmarkdown` syntax. While the initial guide covers basic syntax, here we further your understanding by covering:

- embedding regression results directly into `LaTeX`-formatted equations.

# Convert Regression Models Directly into Latex Equations

The `equatiomatic` package provides an elegant solution for displaying `LaTeX`-formatted equations derived from models specified in the report's R code. Let's delve into its capabilities with practical examples:

### A Simple Regression Model {-}
Transform a model into `LaTeX`, by passing your model inside `extract_eq()`:

```{r}
# To install equatiomatic, you can use: 
## remotes::install_github("datalorax/equatiomatic")
## install.packages("equatiomatic")

library(equatiomatic)

# For a basic model:
fit <- lm(mpg ~ cyl + disp, mtcars)
# Display the theoretical model as an equation
equatiomatic::extract_eq(fit)

# For the model with actual coefficients:
equatiomatic::extract_eq(fit, use_coefs = TRUE)
```

### Categorical Variables {-}

The `equatiomatic` package supports the automatic conversion of categorical variables to `LaTeX` formatted equations. When using categorical variables, it will recognize and automatically include the levels of these variables as subscripted variables in the equation.

```{r, echo=FALSE}
fit_cat <- lm(Sepal.Length ~ Species, iris)
equatiomatic::extract_eq(fit_cat)
```

### Interaction terms {-}

The package also supports the conversion of interaction terms to `LaTeX` formatted equations.

```{r}
fit_int <- lm(data = iris, Sepal.Length ~ Species * Petal.Length)
equatiomatic::extract_eq(fit_int)
```

### Customizing the Output {-}

Adjust the equation output using the following options: 
- `use_coefs`: This option reveals the actual coefficients within the model. 
  - Default = FALSE
- `terms_per_line`: Organizes the equation to display a specific number of terms on the equation's right-hand side. 
- `wrap`: Ensures the equation fits within a predetermined width.

```{r, warning= FALSE, message = FALSE}
library(palmerpenguins)
fit_glm <- glm(species ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = penguins, family = binomial)
equatiomatic::extract_eq(fit_glm, use_coefs = TRUE, terms_per_line = 3, wrap = 80)

fit_probit <- glm(species ~ bill_length_mm + bill_depth_mm + flipper_length_mm, data = penguins, family = binomial(link = "probit"))
equatiomatic::extract_eq(fit_probit, use_coefs = FALSE, terms_per_line = 2, wrap = 40)
```