---
title: "Modelsummary building block"
author: "Valerie Vossen"
date: "2023-04-04"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

# Overview

The modelsummary package is a powerful and user-friendly package for summarizing regression results in R. It allows for easy comparison of multiple models, and offers a wide range of options for controlling the output format. This page will provide you with a summary of the most useful functions by using an example table. The example will be Table 1 of "Doing well by doing good?: Green office buildings" by Eiccholtz et al. (2010). 

To apply the functions of modelsummary, step 1 will be to replicate the regression results of Table 1 of Eiccholtz et al. After that, functions of modelsummary are applied to improve the output table step-by-step.

Steps:
  1. Replicate Table 1 of Eiccholtz et al.
  2. Report different kind of standard errors
  3. Selecting and formatting estimates, statistics, title, and notes
  4. Add stars
  5. Change number of decimals
  6. Report confidence interval instead of standard error
  7. Exporting the output


## Step 1: Replicate Table 1 of Eiccholtz et al. 

Eiccholtz et al. (2010) estimate the effect of environmentally sustainable building practices upon economic outcomes. Table 1 provides us with the estimation outcomes of the regression of a green rating on rental rates for office buildings. There are two kinds of green labels, namely Energy Star and LEED. In the following section, Table 1 will be replicated. 


### Load packages and data 
```{r warning = FALSE, message = FALSE}

# Load packages
library(modelsummary)
library(dplyr)
library(fixest)
library(stringr)

# Load data

data_rent_100710 <-   read.delim("C:/Users/valer/OneDrive/TSH werk/modelsummaryexample/data_rent_100710.txt")

data_rent <- data_rent_100710 %>%
  janitor::clean_names() #clean column names; for example, it converts all letters to lowercase

```

### Modify the data set

```{r}
#Employment growth variable
data_rent$empl_gr <- as.numeric(gsub("%", "",data_rent_100710$empl_gr))     
        #remove % and change into numeric

#Dummy's for stories & age
data_rent$story_high <- ifelse(data_rent$stories>20 & !is.na(data_rent$stories), 1, 0)
data_rent$story_medium <- ifelse(data_rent$stories<21 & data_rent$stories>10 & !is.na(data_rent$stories), 1, 0)
data_rent$story_low <- ifelse(data_rent$stories<11 & !is.na(data_rent$stories), 1, 0)
data_rent$story_low <- ifelse(is.na(data_rent$stories), 1, 0)

data_rent$age_0_10 <- ifelse(data_rent$age < 11 & !is.na(data_rent$age) , 1, 0)
data_rent$age_10_20 <- ifelse(data_rent$age > 10 & data_rent$age < 21 & !is.na(data_rent$age) , 1, 0)
data_rent$age_20_30 <- ifelse(data_rent$age > 20 & data_rent$age < 31 & !is.na(data_rent$age) , 1, 0)
data_rent$age_30_40 <- ifelse(data_rent$age > 30 & data_rent$age < 41 & !is.na(data_rent$age) , 1, 0)
data_rent$age_40 <- ifelse(data_rent$age > 40 & !is.na(data_rent$age) , 1, 0)
data_rent$age_40 <- ifelse(is.na(data_rent$age), 1, 0)

#To deal with NA values, put them in the last category, so no observations get removed because of NA values

#Other changes

data_rent$size_new <- data_rent$size/1000000
  #Size is displayed in millions of square feet: so divide by a million

data_rent$oocc_new <- data_rent$leasing_rate/100
  #fraction occupied: leasing rate in %

data_rent$empl_new <- data_rent$empl_gr/100
  #fraction variable so divided by 100

data_rent$logrent <- log(data_rent$rent)
  #dependent variable is in log

```

### Regression equations

Below, we are estimating regression 1 until 5 displayed in Table 1 of Eiccholtz et al. (2010). To control for locational effects, each regression also includes 694 dummy variables, one for each locational cluster. Regression (5) also includes an additional 694 dummy variables, one for each green building in the sample

```{r warning=FALSE, message=FALSE}

reg1 <- feols(logrent ~ green_rating + size_new + oocc_new + class_a + class_b + net + empl_new| id, data = data_rent)

reg2 <- feols(logrent ~ energystar + leed + size_new + oocc_new + class_a + class_b + net + empl_new| id, data = data_rent)

reg3 <- feols(logrent ~ green_rating + size_new + oocc_new + class_a + class_b + net + empl_new + age_0_10 + age_10_20 + age_20_30 + age_30_40 + renovated | id, data = data_rent)

reg4 <- feols(logrent ~ green_rating + size_new + oocc_new + class_a + class_b + net + empl_new + age_0_10 + age_10_20 + age_20_30 + age_30_40 + renovated + story_medium + story_high + amenities  | id, data = data_rent)
  #regression 1 until 4 include fixed effects for "id"

reg5 <- feols(logrent ~ size_new + oocc_new + class_a + class_b + net + empl_new  + renovated + + age_0_10 + age_10_20 + age_20_30 + age_30_40 + story_medium + story_high + amenities | id + green_rating, data = data_rent)
  #regression 5 includes fixed effects for "id" and "green_rating" variable

```

#### Notes about regression equations:
- 78 observations are removed because of NA values in all 5 regressions. This results in a similar number of observations of 8105 as reported in Table 1 of Eiccholtz et al. (2010). 

- The variable "empl_new" is removed because of collinearity. "empl_new" defines the increase in employment in the service sector for the CBSA containing a cluster of a green building and its nearby control. empl_new value is the same for all variables in one cluster; therefore it is already includes in the fixed effects "id" and it makes sense that R removes the variable. However, Eiccholtz et al. (2010) do report an estimate for "empl_new"

- There is no "constant" in our output.Fixed effects models estimate separate intercepts for each level of a categorical variable (each id), which means that including a global intercept term would lead to perfect multicollinearity. Here, the model estimates a separate intercept term for each unique value of id, so there is no need to include a global intercept term. However, Eiccholtz et al. (2010) do report a constant. 

### Output of table with modelsummary

Now we have estimated the regression equations for our table, we can move on to applying modelsummary! "models" defines a list of regression 1 until 5. Just using modelsummary() without giving any extra input than the models gives us the following output. 

From here, we will work our way through the different functions and improve the table on the way!

```{r warning=FALSE}

models <- list(
  "(1)" = reg1, 
  "(2)" = reg2, 
  "(3)" = reg3, 
  "(4)" = reg4, 
  "(5)" = reg5)

msummary(models)
```


## Step 2: Report different kind of standard errors

Within the modelsummary function, it is possible to specify different types of standard errors, such as robust or clustered standard errors. To replicate the output of Eiccholtz et al. (2010), we use robust standard errors by specifying "vcov = "HC1"". This is appropriate when the variance of the errors varies across observations, which can result in biased standard errors.

On the other hand, clustered standard errors are appropriate when there are groups of observations that are likely to be correlated with each other. To specify standard errors clustered around the id values, we could use "cluster = ~id". Note that in the previous table (in step 1), the default behavior is to cluster standard errors by the variable used to estimate the fixed effects, which in this case is the id variable.

Since our regression models include fixed effects, using robust standard errors is sufficient to account for heteroskedasticity and produce standard errors that are consistent and unbiased. 

```{r}
#robust standard errors
msummary(models, vcov = "HC1")
```

## Step 3: Selecting and formatting estimates, statistics, title, and notes"

This step shows you how to use customization options and create a table that closely matches the formatting of the original Table 1. 

- "coef_map = cm" is used to specify the variable names in our output table, and to rearrange the order of the rows to match Table 1.

- "gof_omit =" allows us to exclude certain statistics from being displayed in the table's bottom section.

- "gof_map = gm" is used to specify which statistics we want to include in the bottom section, as well as their desired formatting. 
The names of the statistics are specified using "clean=" and the number of decimals places are specified using "fmt=".

- We also add a title and note to our table.

```{r}

cm = c('green_rating'    = 'Green rating (1 = yes)',
  'energystar'    = 'Energystar (1 = yes)',
  'leed'    = 'LEED (1 = yes)',
  'size_new'    = 'Building size (millions of sq.ft)',
  'oocc_new' = 'Fraction occupied',
  'class_a' = 'Building class A (1 = yes)',
  'class_b' = 'Building class B (1 = yes)',
  'net' = 'Net contract (1 = yes)',
  'age_0_10' = 'Age: <10 years',
  'age_10_20' = 'Age: 10-20 years',
  'age_20_30' = 'Age: 20-30 years',
  'age_30_40' = 'Age: 30-40 years',
  'renovated' = 'Renovated (1 = yes)',
  'story_medium' = 'Stories: Intermediate (1 = yes)', 
  'story_high' = 'Stories: High (1 = yes)', 
  'amenities' = 'Amenities (1 = yes)')

gm <- list(
  list("raw" = "nobs", "clean" = "Sample size", "fmt" = 0),
  list("raw" = "r.squared", "clean" = "R<sup>2</sup>", "fmt" = 2),
  list("raw" = "adj.r.squared", "clean" = "Adjusted R<sup>2</sup>", "fmt" = 2))
      #get_gof(reg1) to see "raw" names of these statistics.


notetable1 <- c(
  "Notes: Each regression also includes 694 dummy variables, one for each locational cluster. Regression (5) also includes an additional 694 dummy variables, one for each green building in the sample. Standard errors are in brackets.")

titletable1 <- 'Table 1—Regression Results, Commercial Office Rents and Green Ratings 
(dependent variable: logarithm of effective rent in dollars per square foot)'

msummary(models, vcov="HC1",
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm, 
         notes = notetable1, 
         title = titletable1)

```

## Step 4: Add stars

The stars in a regression table are used to indicate the level of statistical significance of the coefficients in the regression model. They are based on the p-values, which measure the probability of obtaining the observed results when there is in fact no effect.

### Argument "stars = TRUE"
To add stars to our regression table, we can use the "stars = TRUE" argument. This will automatically add stars to the table based on a default threshold of statistical significance. By default, a note explaining the significance levels will be added at the bottom of the table: "+ p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001". 

```{r}
msummary(models,
         vcov = "HC1",
         stars=TRUE,
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm, 
         notes = notetable1, 
         title = titletable1)
```

### Manually add stars

To replicate Table 1 of Eiccholtz et al. (2010), we need to customize the output of the regression table to show the significance of coefficients using stars. 

- By default, the stars are printed next to the coefficient estimate, but we want the stars to be printed on the row of the standard error. This can be done by manually adding a list for stars and adding it to the "statistics = " argument. Check the code block below!
- Additionally, we want a different note about the stars so we change our note and add this to the "notes =" argument. 

```{r}

note2table1 <- c(
  "Notes: Each regression also includes 694 dummy variables, one for each locational cluster. Regression (5) also includes an additional 694 dummy variables, one for each green building in the sample. Standard errors are in brackets.",
  "***Significant at the 1 percent level.", 
  "**Significant at the 5 percent level.",
  "*Significant at the 10 percent level.")


msummary(models,
         stars  = c('*' = .1, '**' = 0.05, '***' = .01),
         estimate = "{estimate}",
         statistic = "[{std.error}]{stars}",
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm,
         notes = note2table1,
         title = titletable1)

```


## Step 5: Change number of decimals.

The "fmt=" argument in the modelsummary functions allows us to control how numeric values are rounded and presented in the table.
In order to match the formatting of Table 1 in Eiccholtz et al. (2010), we need to set the number of decimal digits to 3. 

There is various ways to get the desired number of decimals. For example, it is possible to give statistics a different number of decimals than estimates, or to display the values in scientific (exponential) notation by specifying fmt = fmt_sprintf("%.3e"). 

```{r}
msummary(models,
         vcov = "HC1",
         fmt = fmt_statistic(estimate = 3, std.error = 3),
        #just adding "fmt = 3" gives same result, since estimate and standard error are both set to 3 decimals.
         stars  = c('*' = .1, '**' = 0.05, '***' = .01),
         estimate = "{estimate}",
         statistic = "[{std.error}]{stars}",
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm,
         notes = note2table1,
         title = titletable1)

```


## Step 6: Report Confidence Interval instead of Standard Error

While using stars as a threshold for significance can be helpful, it can also be arbitrary and may not always reflect the true importance of a coefficient. Therefore, it is often better to report confidence intervals (CI) in addition to stars. This provides readers with more information about the precision and uncertainty of the estimated effect or parameter.

The CI provides a range of plausible values for the true population parameter. Also when an estimated effect is not statistically significant, the CI gives relevant information about the estimate rather than simply concluding that there is no effect! In this step, we will change the output table to print the CI of the estimate instead of the standard error.

```{r}

msummary(models,
         vcov = "HC1",
         fmt = fmt_statistic(estimate = 3, conf.int = 3),
         statistic ='conf.int',
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm,
         notes = note2table1,
         title = titletable1)

#We can still show stars in the output. (Another way to print the CI is including the lower and upper bound of the CI separately.)

msummary(models,
         vcov = "HC1",
         fmt = fmt_statistic(estimate = 3, std.error = 3),
         stars  = c('*' = .1, '**' = 0.05, '***' = .01),
         estimate = "{estimate}",
         statistic = "[{conf.low}, {conf.high}]{stars}",
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|Std.Errors|FE',
         gof_map = gm,
         notes = note2table1,
         title = titletable1)


```


## Step 7: Exporting the table

The 'output =' argument specifies the destination where the output should be printed. In the given code, the output is named "table1.html". The extension ".html" specifies that the table will be printed in HTML format. Other possible extensions are for example ".docx", ".md", and ".txt". 

By default, if the 'output' argument is not included, the output will be printed directly in the console. 

```{r}

msummary(models, output = "table.html",
         vcov = "HC1",
         fmt = fmt_statistic(estimate = 3, std.error = 3),
         stars  = c('*' = .1, '**' = 0.05, '***' = .01),
         estimate = "{estimate}",
         statistic = "[{std.error}]{stars}",
         coef_map = cm, 
         gof_omit = 'AIC|BIC|RMSE|Within|FE',
         gof_map = gm,
         notes = note2table1,
         title = titletable1)

```

# Summary 

In this building block, we covered the most useful functions of the modelsummary package:

- You can customize the standard errors printed in your model. For instance, "vcov = "HC1"" will give robust standard errors, and "cluster= ~id" will produce clustered standard errors by the variable "id".

- You can customize the way estimates and statistics are presented, including the order of the estimates, variable names and which goodness-of-fit measures are printed. Additionally, you can choose to include confidence intervals instead of standard errors for your estimates. It is also possible to add a title and note at the bottom of your table. 

- You can add stars to indicate the level of statistical significance of the coefficients with "stars = TRUE", which provides default values and a note at the bottom. Alternatively, you can add stars manually to customize the threshold for significance or the note at the bottom.

- The "fmt" argument allows you to specify the number of decimal places for estimates and statistics. 

- You can specify where the table should be printed using the "output" argument. Adding extensions like ".html" and ".docx" will produce the table in that format. 

